Comparison of dimensionality reduction schemes for direct global optimization algorithms

### 2018-05-10 2018/YSC/ysc_paper.tex
YSC: delete parallel sections

%% Front Matter
%%
% Regular title as in the article class.
%
\title{Comparison of dimensionality reduction schemes for parallel global optimization
algorithms}

@@

\titlerunning{Dimension reduction schemes in parallel GO}

% Authors are joined by \and. Their affiliations are given by \inst, which indexes into the list
% defined using \institute
%
\author {
Vladislav Sovrasov \and
Ilya Lebedev
}

% Institutes for affiliations are also joined by \and,
\institute  {
  Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia
  \email{sovrasov.vlad@gmail.com}\\
  \email{ilya.lebedev@itmm.unn.ru}\\
}
%  \authorrunning{} has to be set for the shorter version of the authors' names;
% otherwise a warning will be rendered in the running heads. When processed by
% EasyChair, this command is mandatory: a document without \authorrunning
% will be rejected by EasyChair

\authorrunning{Sovrasov V., Lebedev I.}

@@

\keywords{Global optimization, Dimension reduction, Parallel algorithms,
Multidimensional multiextremal optimization, Global search algorithms}

@@

%------------------------------------------------------------------------------
\subsection{Non-Univalent evolvent}
%\begin{Russian}

@@

A potentially large number of preimages (up to $2^N$) and the inability to use the parallel
scheme for the multiple mappings form Sec.~\ref{sec:parallel_evolvents} are the disadvantages
of the non-univalent evolvent.

@@

%\end{Russian}

@@

\section{Parallel Computations for Solving Global Optimization Problems.}
\subsection{Core Multidimensional Algorithm of Global Search (MAGS)}

@@

%------------------------------------------------------------------------------
\subsection{Parallel algorithm exploiting a set of evolvents}
\label{sec:parallel_evolvents}
Using the multiple mapping allows solving initial problem (\ref{eq:task}) by parallel solving the
problems
\[
\min\{\varphi(y^s(x)):x\in [0,1]\}, 1\leqslant s\leqslant S
\]
on a set of intervals $[0,1]$ by the index method. Each one-dimensional problem is solved on a
separate processor. The trial results at the point \(x^k\) obtained for the problem being solved by
particular processor are interpreted as the results of the trials in the rest problems (in the
corresponding points \(x^(k_1 ),\dots,x^(k_S))\). In this approach, a trial at the point \(x^k \in
[0,1]\) executed in the framework of the \(s\)-th problem, consists in the following sequence of
operations.
\par
1. Determine the image \(y^k=y^s (x^k)\) for the evolvent \(y^s (x)\).
\par
2. Inform the rest of processors about the start of the trial execution at the point\( y^k\) (the
blocking of the point \(y^k\) ).
\par
3. Determine the preimages \(x{}^{k_s}  \in [0,1], 1\leqslant s\leqslant S\), of the point \(y^k\) and interpret the
trial executed at the point \(y^k \in D \) as the execution of the trials in the \(S\) points
\(x{}^{k_1} ,\dots,x{}^{k_s} \)
\par
4. Inform the rest of processors about the trial results at the point \(y^k\).
\par

The decision rules for the proposed parallel algorithm, in general, are the same as the rules of the
sequential algorithm (except the method of the trial execution). Each processor has its own copy
of the software realizing the computations of the problem functions and the decision rule of the
index algorithm. For the organization of the interactions among the processors, the queues are
created on each processor, where the processors store the information on the executed iterations
in the form of the tuples: the processor number \(s\), the trial point \(x{}^{k_s}\).
\par
The proposed parallelization scheme was implemented with the use of MPI technology. Main
features of implementation consist in the following. A separate MPI-process is created for each
of \(S\) one-dimensional problems being solved, usually, one process per one processor
employed. Each process can use $p$ threads, usually one thread per an accessible core.
\par
At every iteration of the method, the process with the index \(s,0\leqslant s< S\) performs p
trials in parallel at the points \(x^{(s+iS)},0\leqslant i<p\). At that, each process stores all
\(S_p\) points, and an attribute indicating whether this point is blocked by another process or
not is stored for each point. Let us remind that the point is blocked if the process starts the
execution of a trial at this point.
\par
At every iteration of the algorithm, operating within the \(s\)-th process, determines the
coordinates of \(p\) ``its own" trial points. Then, the interchange of the coordinates of images of
the trial points \(y^{(s+iS)},0\leqslant i<p, 0\leqslant s< S\) is performed (from each process to
each one). After that, the preimages \(x^{(q+iS)},0\leqslant q<S,q\not=s\) of the points
received by the \(s\)-th process from the neighbor ones are determined with the use of the
evolvent \(y^s (x)\). The points blocked within the \(s\)-th process will correspond to the
preimages obtained. Then, each process performs the trials at the non-blocked points, the
computations are performed in parallel using OpenMP. The results of the executed trials (the
index of the point, the computed values of the problem functions, and the attribute of
unblocking of this point) are transferred to all rest processes. All the points are added to the
search information database, and the transition to the next iteration is performed.

@@

The computational experiments have been carried out on the Lobachevsky supercomputer at
State University of Nizhni Novgorod. A computational node included 2 Intel
Sandy Bridge E5-2660 2.2 GHz processors, 64 GB RAM. The CPUs had 8 cores (i. e. total 16
cores were available per a node). All considered algorithms and evolvents were implemented
using C++ within the Globalizer software system~\cite{globalizerSystem}.
In order to enable the parallelism,
OpenMP was used on a single node, and MPI was used for the parallelization on
several nodes.

>>>>

%% Front Matter
%%
% Regular title as in the article class.
%
\title{Comparison of dimensionality reduction schemes for direct global optimization
algorithms}

@@

\titlerunning{Dimension reduction schemes in direct GO}

% Authors are joined by \and. Their affiliations are given by \inst, which indexes into the list
% defined using \institute
%
\author {
Vladislav Sovrasov
}

% Institutes for affiliations are also joined by \and,
\institute  {
  Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia
  \email{sovrasov.vlad@gmail.com}\\
}
%  \authorrunning{} has to be set for the shorter version of the authors' names;
% otherwise a warning will be rendered in the running heads. When processed by
% EasyChair, this command is mandatory: a document without \authorrunning
% will be rejected by EasyChair

\authorrunning{Sovrasov V.}

@@

\keywords{Global optimization, Dimension reduction,
Multidimensional multiextremal optimization, Global search algorithms}

@@

%------------------------------------------------------------------------------
\subsection{Non-Univalent evolvent}

@@

A potentially large number of preimages (up to $2^N$) and the inability to use the parallel
scheme for the multiple mappings are the disadvantages of the non-univalent evolvent.

@@

@@

\section{Core Multidimensional Algorithm of Global Search}

@@

@@

The computational experiments have been carried out on the Lobachevsky supercomputer at
State University of Nizhni Novgorod. A computational node included 2 Intel
Sandy Bridge E5-2660 2.2 GHz processors, 64 GB RAM. The CPUs had 8 cores (i. e. total 16
cores were available per a node). All considered algorithms and evolvents were implemented
using C++ within the Globalizer software system~\cite{globalizerSystem}.
